taxadb: A High-Performance Local Taxonomic Database Interface

### 2021-03-04 paper/manuscript.Rmd

---
title: "taxadb: A High-Performance Local Taxonomic Database Interface"
author:
  - name: "Kari E A Norman"
    affiliation: ucb
  - name: "Scott Chamberlain"
    affiliation: ropensci
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |
  A familiar and growing challenge in ecological and evolutionary research is that of establishing consistent taxonomy when combining data from separate sources. While this problem is already well understood and numerous naming authorities have been created to address the issue, most researchers lack a fast, consistent, and intuitive way to retrieve taxonomic names.  Here, we present `taxadb` R package to address this gap.  In contrast to existing tools, `taxadb` provides all of the following in one place:  
  1) `taxadb` accesses established naming authorities to resolve synonyms, taxonomic identifiers, and hierarchical classification.  
  2) `taxadb` creates a local database, managed automatically from within R, that provides fast operations on millions of taxonomic names.  
  3) `taxadb` provides a consistent and intuitive data format.  
  4) `taxadb` is built on a simple, extensible, and language-agnostic design that can easily accommodate new authorities.

>>>>

---
title: "taxadb: A High-Performance Local Taxonomic Database Interface"
author:
  - name: "Kari E A Norman"
    affiliation: ucb
  - name: "Scott Chamberlain"
    affiliation: ropensci
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |
  1)  A familiar and growing challenge in ecological and evolutionary research is that of establishing consistent taxonomy when combining data from separate sources. While this problem is already well understood and numerous naming authorities have been created to address the issue, most researchers lack a fast, consistent, and intuitive way to retrieve taxonomic names. 
  2) We present `taxadb` R package which creates a local database, managed automatically from within R, to provide fast operations on millions of taxonomic names. 
  3) `taxadb` provides access to established naming authorities to resolve synonyms, taxonomic identifiers, and hierarchical classification in a consistent and intuitive data format.
  4) `taxadb` makes operation on millions of taxonomic names fast and manageable.
